Supporting Urban Search \& Rescue Mission Planning\\ through Visualization-Based Analysis

### 2014-09-24 vmv14-usar.tex
vmv14-usar: commited camera-ready version
git-svn-id: https://scivis.itn.liu.se/svn/publications/workinprogress/vmv14-usar@5793 cd0c9ba3-fe29-0410-b0c2-9a6eeba77867

% --- for  Annual CONFERENCE
% \ConferenceSubmission % uncomment for Conference submission
% \ConferencePaper      % uncomment for (final) Conference Paper
% \STAR                 % uncomment for STAR contribution
% \Tutorial             % uncomment for Tutorial contribution
% \ShortPresentation    % uncomment for (final) Short Conference Presentation
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to Computer Graphics Forum, special issue
% \SpecialIssuePaper         % uncomment for final version of Journal Paper, special issue
%
% --- for  EG Workshop Proceedings
 \WsSubmission    % uncomment for submission to EG Workshop
% \WsPaper         % uncomment for final version of EG Workshop contribution
%
 \electronicVersion % can be used both for the printed and electronic version

@@

\author{paper1006}
%\author[Bock \etal]{
%    Alexander Bock$^1$, Alexander Kleiner$^2$, Jonas Lundberg$^3$, and Timo Ropinski$^1$\\%
%    $^1$Scientific Visualization Group, Link\"oping University\\%
%    $^2$Collaborative Robotics Group, Link\"oping University\\%
%    {\small $^3$Graphic Design Group, Link\"oping University}
%}

@@

Structural damage to inhabited buildings is an ever-present danger. As time-to-rescue is an important factors for victims' survivability, it is of highest importance for rescue responders to reach victims and provide medical attention or extraction. Planning and executing paths through buildings, however, is difficult as available floor plans are outdated. Furthermore, structural weaknesses as well as hazardous environments, make regions and areas inaccessible. While trained responders can use their knowledge and intuition to spot these hazards, it is paramount that the \emph{Incident Commander} (IC) can analyze the information and coordinate rescue responders. There exist well-defined protocols describing the actions taken during an \emph{Urban Search~\&~Rescue} (USAR) operation. In most protocols, one \IC\ is responsible for a single building and instructs multiple rescue responders inside. The Federal Emergency Management Agency describes steps that are performed by the rescue team~\cite{fema08}. During the assessment step, 2D maps of the collapsed building are drawn based on the descriptions of rescue responders moving about the building searching for victims. In this exploration phase, they might stumble into hazardous areas that endanger the rescuer's life.

@@

\section{Decision-Making Theory} \label{sec:theory}
It is crucial to take knowledge about human decision making into account to design a decision support system. Decision makers tend to evaluate options serially in time-constrained situations; they attempt to find one viable plan rather than attempting to generate and compare numerous plans in parallel. This theory has been described by Klein and Calderwood as \emph{Recognition Primed Decision-making} (RPD)~\cite{KleinCalderwood}. Initially, experts look for similarities to previous situations with regard to relevant goals, things that were important to monitor, and possible actions. Then, they go through a process of mental simulation to consider whether the actions are applicable to the case at hand. They assess the ongoing situation, looking for violations and confirmations of expectancies, which may require reframing the situation. Klein and Calderwood suggest that ``displays and interfaces should be centered on decisions rather than around data flows''~\cite{KleinCalderwood}, emphasizing that systems can be built to enhance the workflow of mental simulation.

@@

\begin{figure}
	\centering
	\includegraphics[height=0.8125\columnwidth]{figures/workflow.pdf}
	\caption{A schematic timeline overview of the currently employed workflow and our proposed system, showing events (red) and actions (blue) split up into five distinct phases. Utilizing the additional actions (yellow) in parallel enables faster exploration and decreased overall time-to-rescue.}
	\label{fig:workflow:workflow}
\end{figure}

@@

\subsection{Data Annotation} \label{sec:overview:annotation}
Each voxel can belong to one of five classes. \emph{Unclassified} is the default type for all voxels. \emph{Start} voxels are accessible entry points. This means that paths can start from any of these points. \emph{POI} voxels are destination points for paths. These indicates that there is a potential victim or another mission-critical element at this location. \emph{Hazard} voxels have been declared as dangerous. Each hazard area has a normalized severity denoting how dangerous it is. \emph{Forbidden} voxels can only be declared by the \IC\ and are areas completely out of reach. These points are used when, for example, a corridor collapses and is not accessible anymore. The classification for each voxel is modified by directly interacting with the 3D rendering, thus fulfilling {\bfseries R2}.

@@

\begin{figure*}
	\centering
	\subfigure[Profile Plot]{
	    \fbox{\includegraphics[width=0.31\textwidth]{figures/fig-analysis-profile.png}}
	    \label{fig:overview:analysis:profile}
	}
	\hfill
	\subfigure[Parallel Coordinates Plot]{
	    \fbox{\includegraphics[width=0.31\textwidth]{figures/fig-analysis-pcp.png}}
	    \label{fig:overview:analysis:pcp}
	}
	\hfill
	\subfigure[Scatter Plot Matrix]{
	    \fbox{\includegraphics[width=0.31\textwidth]{figures/fig-analysis-scatter.jpg}}
	    \label{fig:overview:analysis:scatter}
	}
	\caption{Analysis views supporting comparative path analysis. (a) the Profile Plot presents the change of a single attribute along each path; here the distance to the closest hazard. (b) the Parallel Coordinates Plot shows correlations between attributes. (c) the Scatter Plot Matrix depicts the correlations of all attributes.}
\end{figure*}

@@

We performed an evaluation of our system with nine external experts working in the field of Urban Search \& Rescue. Of these experts, five (A--E) are emergency responders, three (F--H) are researchers, and one (J) is a consultant for a national technical relief agency. As the goal of the study was to include as many international experts as possible, and being aware of their time constraints, we decided to create videos and images of our system for the application case and created an interactive webpage rather than requiring training for hands-on use of our system. 7 of 9 experts finished the evaluation and provided answers to all questions. In the analysis, we consider the partial responses for the cases where an answer was provided. The full evaluation and replies are available in the supplemental material. Here, we will discuss important positive and negative answers by the experts. \\

@@

For future work, we like to perform a thorough in-use evaluation workshop with domain experts accompanying a real-world rescue scenario. This will provide us with valuable direct feedback to improve usability of the system. This includes researching methods on how to familiarize the SAR experts with the PCP and the SPLOM method. Furthermore, we plan to investigate the applicability of adaptive and stochastic sampling, like Monte-Carlo methods, on the high-dimensional parameter space for path computation.

@@

%\section*{Acknowledgments}
%We like to thank the IRIDeS, the CREATE, and the IRS institutes for their cooperation in retrieving the Tohoku university scans.
%This work was partly supported by grants from the Excellence Center at Link\"oping and Lund in Information Technology (ELLIIT) and the Swedish e-Science Research Centre (SeRC), as well as VR grant 2011-4113. We would like to thank the external experts for taking their time to take part in the evaluation. The presented concepts have been realized using the Voreen framework (www.voreen.org).

>>>>

% --- for  Annual CONFERENCE
% \ConferenceSubmission % uncomment for Conference submission
% \ConferencePaper      % uncomment for (final) Conference Paper
% \STAR                 % uncomment for STAR contribution
% \Tutorial             % uncomment for Tutorial contribution
% \ShortPresentation    % uncomment for (final) Short Conference Presentation
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to Computer Graphics Forum, special issue
% \SpecialIssuePaper         % uncomment for final version of Journal Paper, special issue
%
% --- for  EG Workshop Proceedings
% \WsSubmission    % uncomment for submission to EG Workshop
 \WsPaper         % uncomment for final version of EG Workshop contribution
%
 \electronicVersion % can be used both for the printed and electronic version

@@

%\author{paper1006}
\author[Bock \etal]{
    Alexander Bock$^1$, Alexander Kleiner$^2$, Jonas Lundberg$^3$, and Timo Ropinski$^1$\\%
    $^1$ Scientific Visualization Group, Link\"oping University\\%
    $^2$ iRobot, Pasadena, CA\\%
    {\small $^3$ Graphic Design Group, Link\"oping University}
}

@@

Structural damage to inhabited buildings is an ever-present danger. As time-to-rescue is an important factors for victims' survivability, it is of highest importance for rescue responders to reach victims and provide medical attention or extraction. Planning and executing paths through buildings, however, is difficult as available floor plans are outdated. Furthermore, structural weaknesses as well as hazardous environments, make regions and areas inaccessible. While trained responders can use their knowledge and intuition to spot these hazards, it is paramount that the \emph{Incident Commander} (IC) can analyze the information and coordinate rescue responders. There exist well-defined protocols describing the actions taken during an \emph{Urban Search~\&~Rescue} (USAR) operation. In most protocols, one \IC\ is responsible for a single building and instructs multiple rescue responders inside. The Federal Emergency Management Agency describes steps that are performed by the rescue team. During the assessment step, 2D maps of the collapsed building are drawn based on the descriptions of rescue responders moving about the building searching for victims. In this exploration phase, they might stumble into hazardous areas that endanger the rescuer's life.

@@

\section{Decision-Making Theory} \label{sec:theory}
It is crucial to take knowledge about human decision making into account to design a decision support system. Andrienko and Andrienki showed that the usage of explorative interactive tools support the decision process~\cite{Andrienko:2003kv}. Decision makers tend to evaluate options serially in time-constrained situations; they attempt to find one viable plan rather than attempting to generate and compare numerous plans in parallel. This theory has been described by Klein and Calderwood as \emph{Recognition Primed Decision-making} (RPD)~\cite{KleinCalderwood}. Initially, experts look for similarities to previous situations with regard to relevant goals, things that were important to monitor, and possible actions. Then, they go through a process of mental simulation to consider whether the actions are applicable to the case at hand. They assess the ongoing situation, looking for violations and confirmations of expectancies, which may require reframing the situation. Klein and Calderwood suggest that ``displays and interfaces should be centered on decisions rather than around data flows''~\cite{KleinCalderwood}, emphasizing that systems can be built to enhance the workflow of mental simulation.

@@

\begin{figure}
	\centering
	\includegraphics[height=0.85\columnwidth]{figures/workflow.pdf}
	\caption{A schematic timeline overview of the currently employed workflow and our proposed system, showing events (red) and actions (blue) split up into five distinct phases. Utilizing the additional actions (yellow) in parallel enables faster exploration and decreased overall time-to-rescue.}
	\label{fig:workflow:workflow}
\end{figure}

@@

\subsection{Data Annotation} \label{sec:overview:annotation}
Each voxel can belong to one of five classes. \emph{Unclassified} is the default type for all voxels. \emph{Start} voxels are accessible entry points. This means that paths can start from any of these points. \emph{POI} voxels are destination points for paths. These indicates that there is a potential victim or another mission-critical element at this location. \emph{Hazard} voxels have been declared as dangerous. Each hazard area has a normalized severity denoting how dangerous it is. \emph{Forbidden} voxels can only be declared by the \IC\ and are areas completely out of reach. These points are used when, for example, a corridor collapses and is not accessible anymore. The classification for each voxel is modified by interacting with the 3D rendering using the mouse and keyboard, thus fulfilling {\bfseries R2}.

@@

\begin{figure*}
	\centering
	\subfigure[Profile Plot]{
	    \fbox{\includegraphics[width=0.28\textwidth]{figures/fig-analysis-profile.png}}
	    \label{fig:overview:analysis:profile}
	}
	\hfill
	\subfigure[Parallel Coordinates Plot]{
	    \fbox{\includegraphics[width=0.28\textwidth]{figures/fig-analysis-pcp.png}}
	    \label{fig:overview:analysis:pcp}
	}
	\hfill
	\subfigure[Scatter Plot Matrix]{
	    \fbox{\includegraphics[width=0.28\textwidth]{figures/fig-analysis-scatter.jpg}}
	    \label{fig:overview:analysis:scatter}
	}
	\caption{Analysis views supporting comparative path analysis. (a) the Profile Plot presents the change of a single attribute along each path; here the distance to the closest hazard. (b) the Parallel Coordinates Plot shows correlations between attributes. (c) the Scatter Plot Matrix depicts the correlations of all attributes.}
\end{figure*}

@@

We performed an evaluation of our system with nine external experts working in the field of Urban Search \& Rescue. Of these experts, five (A--E) are emergency responders, three (F--H) are researchers, and one (J) is a consultant for a national technical relief agency. As the goal of the study was to include as many international experts as possible, and being aware of their time constraints, we decided to create videos and images of our system for the application case and created an interactive webpage rather than requiring training for hands-on use of our system. 7 of 9 experts finished the evaluation and provided answers to all questions. In the analysis, we consider the partial responses for the cases where an answer was provided. The full evaluation and replies are available in the supplemental material. Here, we will discuss important positive and negative answers by the experts. The replies to these questions are on a 5-point Likert scale. \\

@@

For future work, we like to perform a thorough in-use evaluation workshop with domain experts accompanying a real-world rescue scenario. This will provide us with valuable direct feedback to improve usability of the system. This includes researching methods on how to familiarize the SAR experts with the PCP and the SPLOM method. Furthermore, we plan to investigate the applicability of adaptive and stochastic sampling, like Monte-Carlo methods, on the high-dimensional parameter space for path computation. In addition, we will investigate the possibility of integrating video footage into the 3D point cloud rendering~\cite{1453517}.

@@

\section*{Acknowledgments}
We like to thank the IRIDeS, the CREATE, and the IRS institutes for retrieving the Tohoku university scans. This work was partly supported by grants from the Excellence Center at Link\"oping and Lund in Information Technology (ELLIIT) and the Swedish e-Science Research Centre (SeRC), as well as VR grant 2011-4113. We thank the external experts for taking part in the evaluation.
